---
title: "fieldClim"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fieldClim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `fieldClim` package was originally designed as a course project to the course `Geländeklimatologie`, held by Prof. Dr. Jörg Bendix at the Philipps-University of Marburg in summer term 2020. Thus, the calculations and formulas of this package are based on this course, as well as the book `Geländeklimatologie` (Field climatology) by Jörg Bendix (2004; ISBN 978-3-443-07139-4).

`fieldClim` is designed as a handy tool, that lets you calculate various weather and micro-climate conditions, based on the measurements of a weather station. It lets you create a `weather-station`-object, that can then be used to call most of the functions without the necessity of further input variables. In addition, all functions can also be called by manually inputting the needed variables, if the user whishes to do so.

Here, we will provide an example of the calculation of a set of micro-climate conditions using the `weather-station`-object, based on the `weather_station_example_data` data frame, that is included in the package.












## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.
